Reagents: expressing and composing patterns of fine-grained concurrency

### 2011-10-12 paper/paper.tex
work on paper intro

``Programs are what happens between cache misses.''

As with most ..., concurrency only exacerbates performance problems
with the memory hierarchy.

ahmdahl's law -- even if you parallelize the big chunks, the
fine-grained stuff will get you in the end

%\section{Reagent semantics}

%\begin{verbatim}

%\end{verbatim}

@@

>>>>

\section{Introduction}

\hspace{\stretch{1}}\emph{Programs are what happens between cache
  misses}.\footnote{Folklore; origin unknown.}

\subsection*{The problem}

Ahmdahl's law tells us that sequential bottlenecks fundamentally limit
profit from parallelism~\cite{?}.  In practice, the effect is made
fatal by a silent killer: interprocessor communication, often in the
form of cache coherence.  When one thread waits on another, the
program pays the cost of lost parallelism \emph{and} an extra cache
miss.  Extra misses can easily add up to parallel \emph{slowdown},
more than negating the benefits of the remaining parallelism.  Cache,
as ever, is king.

The easy answer is: avoid communication.  In other words, parallelize
at a coarse grain, giving threads large chunks of independent work.
That strategy only goes so far.  Some work doesn't easily factor into
large chunks, or equal size chunks.  Fine-grained parallelism is
easier to find, and easier to sprinkle throughout existing sequential
code.  With increasing pressure to parallelize, fine-grained
parallelism will only become more common.

In the past two decades,\footnote{Herlihy/Wing, Mellor-Crummey/Scott}
an enormous literature has emerged

\subsection*{Our contribution}

\section{Reagents}

\section{Semantics}

\section{Implemention}

\section{Performance}

\section{Related work}

@@

# Intro notes

What do we want to accomplish?

 - motivate libraries like j.u.c
 - downsides of such libraries
   - huge research & impl effort
   - not easily extensible
   - internal structure has repeated, but unabstracted patterns
 - our contribution
   - a new way of expressing scalable algorithms that
     - doesn't lose much performance 
     - allows composition
     - is much nicer than direct programming
     - also incorporates blocking behavior
